Category-Level Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection

### 2017-03-26 paper/presentation/DASFAA_presentation.tex
presentation 1.0

Many Web applications need the \textbf{accurate event detection} technique on microblog stream, including:
\begin{enumerate}
	\item public opinion analysis [Chen, SIGIR 2013]
	\item public security [Li, ICDE 2012], [Imran, WWW 2014]
	\item disaster response [Sakaki,WWW2010]
	\item breaking news report\footnote{\url{http://www.theverge.com/2016/12/1/13804542/reuters-algorithm-breaking-news-twitter}}
\end{enumerate}	
\vfill

@@

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB)}	
There is a three fold hierarchical structure in Knowledge Base.
\begin{enumerate}
	\item \textbf{Taxonomy Graph \(G^{(0)}\)}. The directed edges in \(G^{(0)}\) represent the \textit{class}\(\rightarrow\)\textit{subclass} relations in KB. 
		\begin{itemize}
			\item e.g., \textit{Main topic classifications} \(\rightarrow\) \textit{Society}
		\end{itemize}
	\item \textbf{Category-Page Bipartite Graph \(G^{(1)}\)}. The directed edges in \(G^{(1)}\) represent the \textit{class}\(\rightarrow\)\textit{instance} relations in KB.
		\begin{itemize}
			\item e.g., \textit{Military} \(\rightarrow\) \textit{page: Armed Forces}
		\end{itemize}
	\item \textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
		\begin{itemize}
			\item e.g., \textit{page: Armed Forces} \(\rightarrow\) \textit{content(20170325version): ``The armed forces of a country are its government-sponsored defense, fighting forces, and organizations..."}
		\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB)}
\begin{algorithm}[H]
\scriptsize
\caption{Extraction of Category-Level Topics in Knowledge Base}
\label{alg:normalStatesInit}

@@

%\begin{comment}
%------------------------------
%page 9
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB) (1/3)}	
Taking the category \textit{Military} as an example, we extract \textit{Military}'s category-level topic \(\bm{h_{Military}}\).
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.45\columnwidth]{../img/initializationExample.pdf}
        \caption{Extracting Category-Level Topics in Knowledge Base via its three fold hierarchical structure, taking \textit{Military} as an example.}
        \label{fig:hood}
\end{figure}

@@

%------------------------------
%page 10
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (2/3)}	
Transfer KB's Category-Level Topics \(\{\bm{h_c}\}_{c=1}^{K_{KB}}\) into microblogs stream: CTrans-LDA.
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.6\columnwidth]{../model/lda_tikz.pdf}
        \caption{Diagram of CTrans-LDA.}
        \label{fig:hood}
\end{figure}

@@

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (3/3)}	
We use Gibbs Sampling for solving CTrans-LDA.
\begin{itemize}
	\item The initialization probability \(\hat{q}_{k|v}\) makes sure that the learned topics are aligned to the pre-defined category-level topic.
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\begin{scriptsize} 
\begin{equation}
\label{eq:initProbability}
\begin{aligned}
\hat{q}_{k|v}=
\left\{ \begin{aligned}
\frac{\tau_{kv}}{\sum_{k=1}^{K}\tau_{kv}} &,\sum_{k}\tau_{kv}>0 & (a)\\
0&, \sum_{k}\tau_{kv}=0 \ and \ k \leq K_{\bm{KB}} & (b)\\
1/(K-K_{\bm{KB}})&,\sum_{k}\tau_{kv}=0 \ and \ k > K_{\bm{KB}} & (c)
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}
\item Conditional probability in gibbs sampling:
\(p(z_{dn}=k|.)\propto (n^{(d)}_{dk}+\alpha m_k)(n^{(w)}_{kv}+\tau_{kv}+\beta)/(n^{(w)}_{k,.}+\tau_{k,.}+V\beta)\).
\end{itemize}

@@

\begin{frame}
\frametitle{Experimental Results}	
\begin{table}[h]
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Overall Performance on Event Detection}

@@

\begin{frame}
\frametitle{Experimental Results}	
\begin{figure}[h]
	\setlength{\abovecaptionskip}{0.cm}
	\setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{../img/barchartOnBenchmark1.pdf}
        \caption{The relation between the recall and the event size}
        \label{fig:Benchmark1}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Experimental Results}	
\begin{table}
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Events about \textit{military} detected by systems between 2011-07-22 and 2011-07-28}
\label{my-label}
\scalebox{0.6}{
\begin{threeparttable}  
\begin{tabular}{|c|l|l|c|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{Date} & \multirow{2}{*}{Event key words} & \multirow{2}{*}{Representative event tweet} & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Number of \\ event tweet\end{tabular}} & \multicolumn{6}{c|}{Methods\tnote{a}} \\ \cline{5-10} 
 &  &  &  & L & TU & TW & E & B & TD \\ \hline
7/22/11 & \begin{tabular}[c]{@{}l@{}}Norway, Oslo,\\ attacks, bombing\end{tabular} & \begin{tabular}[c]{@{}l@{}}Terror Attacks Devastate Norway: A bomb\\ ripped through government offices in Oslo \\and a gunman... http://dlvr.it/cLbk8\end{tabular} & 557 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
7/23/11 & Gunman, rink & \begin{tabular}[c]{@{}l@{}}Gunman Kills Self, 5 Others at Texas Roller\\ Rink http://dlvr.it/cLcTH\end{tabular} & 43 & - & - & \checkmark &  \checkmark & - & \checkmark \\ \hline
7/26/11 & \begin{tabular}[c]{@{}l@{}}Kandahar, mayor, \\ suicide, attack\end{tabular} & \begin{tabular}[c]{@{}l@{}}TELEGRAPH{]}: Kandahar mayor killed by\\ Afghan suicide bomber: The mayor of \\Kandahar, the biggest city in south \_\end{tabular} & 47 & \checkmark & - & \checkmark & \checkmark & - & \checkmark \\ \hline
7/28/11 & Ft., Hood, attack & \begin{tabular}[c]{@{}l@{}} Possible Ft. Hood Attack Thwarted\\ http://t.co/BSJ33hk\end{tabular} & 52 & - & - & - & - & - & \checkmark \\ \hline
7/28/11 & \begin{tabular}[c]{@{}l@{}}Libyan, rebel, \\ gunned\end{tabular} & \begin{tabular}[c]{@{}l@{}}Libyan rebel chief gunned down in Benghazi \\ http://sns.mx/prfvy1\end{tabular} & 44 & - & - & - & - & - & \checkmark \\ \hline
\end{tabular}

>>>>

Many Web applications need the \textbf{accurate event detection} technique on microblog stream, including:
\begin{enumerate}
	\item public opinion analysis [Chen, SIGIR 2013]
	\item public security [Li, ICDE 2012], [Imran, WWW 2014]
	\item disaster response [Sakaki,WWW 2010]
	\item breaking news report\footnote{\url{http://www.theverge.com/2016/12/1/13804542/reuters-algorithm-breaking-news-twitter}}
\end{enumerate}	
\vfill

@@

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB) (1/3)}	
There is a three fold hierarchical structure in Knowledge Base.
\begin{enumerate}
	\item \textbf{Taxonomy Graph \(G^{(0)}\)}. The directed edges in \(G^{(0)}\) represent the \textit{class}\(\rightarrow\)\textit{subclass} relations in KB. 
		\begin{itemize}
			\item e.g., \textit{Main topic classifications} \(\rightarrow\) \textit{Society}
		\end{itemize}
	\item \textbf{Category-Page Bipartite Graph \(G^{(1)}\)}. The directed edges in \(G^{(1)}\) represent the \textit{class}\(\rightarrow\)\textit{instance} relations in KB.
		\begin{itemize}
			\item e.g., \textit{Military} \(\rightarrow\) \textit{page: Armed Forces}
		\end{itemize}
	\item \textbf{Page-Content Map \(G^{(2)}\)}. 
For a specific Wikipedia dumps version, the edges  \textit{page} \(\rightarrow\)\textit{content} in \(G^{(2)}\) define a one-to-one mapping.
		\begin{itemize}
			\item e.g., \textit{page: Armed Forces} \(\rightarrow\) \textit{content(20170325version): ``The armed forces of a country are its government-sponsored defense, fighting forces, and organizations..."}
		\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB) (2/3)}
\begin{algorithm}[H]
\scriptsize
\caption{Extraction of Category-Level Topics in Knowledge Base}
\label{alg:normalStatesInit}

@@

%\begin{comment}
%------------------------------
%page 9
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 1 (Extracting Category-Level Topics in KB) (3/3)}	
Taking the category \textit{Military} as an example, we extract \textit{Military}'s category-level topic \(\bm{h_{Military}}\).
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.45\columnwidth]{../img/initializationExample.pdf}
        \caption{Extracting Category-Level Topics in Knowledge Base via its three fold hierarchical structure, taking \textit{Military} as an example.}
        \label{fig:hood}
\end{figure}

@@

%------------------------------
%page 10
\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (1/2)}	
Transfer KB's Category-Level Topics \(\{\bm{h_c}\}_{c=1}^{K_{KB}}\) into microblogs stream: CTrans-LDA.
\begin{figure}[h]
		\setlength{\abovecaptionskip}{0.cm}
        \setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=0.6\columnwidth]{../model/lda_tikz.pdf}
        \caption{Diagram of CTrans-LDA.}
        \label{fig:hood}
\end{figure}

@@

\begin{frame}
\frametitle{\textsc{TransDetector}: Phase 2 (Transferring Category-Level Info into Microblog Stream) (2/2)}	
We use Gibbs Sampling for solving CTrans-LDA.
\begin{itemize}
	\item The initialization probability \(\hat{q}_{k|v}\) makes sure that the learned topics are aligned to the pre-defined category-level topic.
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{0pt}
\begin{scriptsize} 
\begin{equation}
\label{eq:initProbability}
\begin{aligned}
\hat{q}_{k|v}=
\left\{ \begin{aligned}
\frac{\tau_{kv}}{\sum_{k=1}^{K}\tau_{kv}} &,\sum_{k}\tau_{kv}>0 & (a)\\
0&, \sum_{k}\tau_{kv}=0 \ and \ k \leq K_{\bm{KB}} & (b)\\
1/(K-K_{\bm{KB}})&,\sum_{k}\tau_{kv}=0 \ and \ k > K_{\bm{KB}} & (c)
\end{aligned}\right.
\end{aligned}
\end{equation}
\end{scriptsize}
\item Conditional probability in gibbs sampling:
\(p(z_{dn}=k|.)\propto (n^{(d)}_{dk}+\alpha m_k)(n^{(w)}_{kv}+\tau_{kv}+\beta)/(n^{(w)}_{k,.}+\tau_{k,.}+V\beta)\).
\end{itemize}

@@

\begin{frame}
\frametitle{Experimental Results}	
Effectiveness of event detection:
\begin{enumerate}
	\item \textsc{TransDetector} performs better in terms of the precision and the recall.
	\item only sacrificing in the DERate slightly because an event could be grouped into multiple categories.
	\begin{itemize}
		\item the event \textit{``S\&P downgrade US credit rating"}, related to \textit{politics} and  \textit{financial} simultaneously.
	\end{itemize}
\end{enumerate}

\begin{table}[h]
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Overall Performance on Event Detection}

@@

\begin{frame}
\frametitle{Experimental Results}
To understand why \textsc{TransDetector} performs better.
\vfill
Show the relationship between the recall and the event size.
\begin{figure}[h]
	\setlength{\abovecaptionskip}{0.cm}
	\setlength{\belowcaptionskip}{0.cm}
        \centering
        \includegraphics[width=1.0\columnwidth]{../img/barchartOnBenchmark1.pdf}
        \caption{The relation between the recall and the event size}
        \label{fig:Benchmark1}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Experimental Results}	
To understand why \textsc{TransDetector} performs better.\\
Show the relationship between the recall and the event size, taking the \textit{military}-related events as an example.
\begin{table}
\setlength{\abovecaptionskip}{0.cm}%set the distance between caption and table to 0 cm.
\setlength{\belowcaptionskip}{0.cm}
\centering
\caption{Events about \textit{military} detected by systems between 2011-07-22 and 2011-07-28}
\label{my-label}
\scalebox{0.6}{
\begin{threeparttable}  
\begin{tabular}{|c|l|l|c|c|c|c|c|c|c|}
\hline
\multirow{2}{*}{Date} & \multirow{2}{*}{Event key words} & \multirow{2}{*}{Representative event tweet} & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Number of \\ event tweet\end{tabular}} & \multicolumn{6}{c|}{Methods\tnote{a}} \\ \cline{5-10} 
 &  &  &  & L & TU & TW & E & B & TD \\ \hline
7/22/11 & \begin{tabular}[c]{@{}l@{}}Norway, Oslo,\\ attacks, bombing\end{tabular} & \begin{tabular}[c]{@{}l@{}}Terror Attacks Devastate Norway: A bomb\\ ripped through government offices in Oslo \\and a gunman... http://dlvr.it/cLbk8\end{tabular} & 557 & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
7/23/11 & Gunman, rink & \begin{tabular}[c]{@{}l@{}}Gunman Kills Self, 5 Others at Texas Roller\\ Rink http://dlvr.it/cLcTH\end{tabular} & 43 & - & - & \checkmark &  \checkmark & - & \checkmark \\ \hline
7/26/11 & \begin{tabular}[c]{@{}l@{}}Kandahar, mayor, \\ suicide, attack\end{tabular} & \begin{tabular}[c]{@{}l@{}}TELEGRAPH{]}: Kandahar mayor killed by\\ Afghan suicide bomber: The mayor of \\Kandahar, the biggest city in south \_\end{tabular} & 47 & \checkmark & - & \checkmark & \checkmark & - & \checkmark \\ \hline
7/28/11 & Ft., Hood, attack & \begin{tabular}[c]{@{}l@{}} Possible Ft. Hood Attack Thwarted\\ http://t.co/BSJ33hk\end{tabular} & 52 & - & - & - & - & - & \checkmark \\ \hline
7/28/11 & \begin{tabular}[c]{@{}l@{}}Libyan, rebel, \\ gunned\end{tabular} & \begin{tabular}[c]{@{}l@{}}Libyan rebel chief gunned down in Benghazi \\ http://sns.mx/prfvy1\end{tabular} & 44 & - & - & - & - & - & \checkmark \\ \hline
\end{tabular}
